---
title: "Join Lab"
author: "Fatima Tanvir"
date: "DATA 2401"
output: html_document
---


# Joins and `dplyr` practice

Load the `"nycflights13"` package.  You'll also need to load `dplyr` or `tidyverse`, if you didn't above.

```{r}
library(tidyverse)
#install.packages("nycflights13")
library(nycflights13)
my_flights <- flights # store my own version
my_flights <- na.omit(my_flights) ## this dataset has data on cancelled flights too, 
                                  ## so we 'll remove those for now

```


Create a dataframe of the average arrival delays for each _destination_

```{r}
average_arrival_delays <- my_flights %>%
  group_by(dest) %>%
  summarise(average_arrival_delay = mean(arr_delay))
average_arrival_delays
```

Then use `left_join()` to join on the "airports" dataframe, which has the airport information.

```{r}
average_arrival_delays_with_info <- average_arrival_delays %>%
  left_join(airports, by = c("dest" = "faa"))
average_arrival_delays_with_info
```


Note here: the column names don't match anymore! 

Which airport had the largest average arrival delay?

```{r}
airport_with_max_arr_delay <- average_arrival_delays_with_info %>%
  arrange(desc(average_arrival_delay)) %>%
  slice_head(n = 1)

airport_with_max_arr_delay

```

What is the average departure delay for each airline? To answer this with something more than just the FAA code of the airline name, you need to create a dataframe of the average departure delay for each _airline_, then use `left_join()` to join on the "airlines" dataframe.


```{r}
average_departure_delays <- my_flights %>%
  group_by(carrier) %>%
  summarise(average_departure_delay = mean(dep_delay))


average_departure_delays_with_name <- average_departure_delays %>%
  left_join(airlines, by = "carrier")

average_departure_delays_with_name
```

Which airline had the smallest average departure delay? 

```{r}
airline_with_min_dep_delay <- average_departure_delays_with_name %>%
  arrange(average_departure_delay) %>%
  slice_head(n = 1)
airline_with_min_dep_delay


hi <- anti_join(flights, airports, by = c("dest" = "faa"))
hi

bye <- anti_join(airports, flights, by = c("faa" = "dest"))
bye
```

Examples from R4DS:    

```{r}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest
```

Now you want to find each flight that went to one of those destinations. You could, of course, filter. But! Now you can semi-join: 

```{r}
flights %>% 
  semi_join(top_dest)
```


What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?
Perform the two joins below and consider what the results are. 
